### 2023-05-02 chapter2/chapter2.tex
Added more citations

This form of selective constraint can also describe the evolution of more ``localized'' features in IDRs such as short linear motifs (SLiMs). Because SLiMs are composed of fewer than 12 residues, they form limited interfaces that frequently mediate the transient binding events involved in signaling pathways~\cite{Tompa2014}. Accordingly, they are highly enriched in IDRs, which provide an accessible and flexible scaffold for these interactions~\cite{Fuxreiter2007, Davey2012}. While some SLiMs in IDRs are strongly conserved at specific positions, these constitute a small fraction of disordered residues, estimated at roughly 17\% in the yeast proteome~\cite{NguyenBa2012}. Instead, as SLiMs are compact and often highly degenerate at some positions, they can arise de novo from a small number of mutations and therefore have high rates of turnover. Furthermore, when IDRs contain multiple copies of a motif that jointly mediate a high-avidity interaction~\cite{Tompa2014} or a graded response to a signal via the accumulation of multiple phosphorylations~\cite{Wright2014}, the individual motifs are under weak selective constraints. As a result, though SLiMs are encoded by specific sequences, in some contexts they may evolve as distributed features that characterize IDRs as a whole rather than specific sites within them.

@@

The GO annotation enrichment analysis indicates that proteins involved in DNA repair and extracellular structures contain a disproportionate number of regions whose disorder scores are rapidly evolving. Because the significance tests were not corrected for multiple testing or controlled for their false discovery rate, we caution against over-interpreting this result and instead consider it as a hypothesis for further investigation. In general, however, these regions with rapidly evolving disorder scores may correspond to molecular recognition features (MoRFs), which are modules in IDRs that undergo a disorder-to-order transition on binding their targets~\cite{vanderLee2014}. Because MoRFs already exist on the boundary between disorder and structure, small changes in the biophysical properties of these regions may have disproportionate effects on their structural ensembles. Thus, the most variable disorder scores may reflect instances where a mutation triggered a ``phase transition'' between largely structured or disordered native states.

@@

\subsection{Brownian motion and Ornstein-Uhlenbeck analyses}
Brownian motion (BM) model parameters were calculated with two methods. The first used Felsenstein's contrasts algorithm to efficiently calculate roots and contrasts for the disorder scores and features of each region~\cite{Felsenstein1973, Felsenstein1985}. Rates were calculated as the mean of the squares of the contrasts. Though these values are unbiased, they are not maximum likelihood estimates and are inappropriate for use with the Akaike information criterion (AIC)~\cite{Akaike1974}. Thus, they were used for analyses involving only the BM model. For comparison with the Ornstein-Uhlenbeck (OU) model, the BM parameters were calculated by maximizing the likelihood. The OU model parameters were also calculated via maximum likelihood estimation. To ensure parameter identfiability and estimate consistency, the root was treated as a random variable~\cite{Ho2013, Ho2014}. Thus, the covariance matrix, $V$, was parameterized as $V_{ij} = e^{\alpha d_{ij}}$ where $d_{ij}$ is the tree distance between tips $i$ and $j$, and $\alpha$ is the selection strength~\cite{Ho2013}.

>>>>

This form of selective constraint can also describe the evolution of more ``localized'' features in IDRs such as short linear motifs (SLiMs). Because SLiMs are composed of fewer than 12 residues, they form limited interfaces that frequently mediate the transient binding events involved in signaling pathways~\cite{Tompa2014}. Accordingly, they are highly enriched in IDRs, which provide an accessible and flexible scaffold for these interactions~\cite{Fuxreiter2007, Davey2012}. While some SLiMs in IDRs are strongly conserved at specific positions, these constitute a small fraction of disordered residues, estimated at roughly 17\% in the yeast proteome~\cite{NguyenBa2012}. Instead, as SLiMs are compact and often highly degenerate at some positions, they can arise de novo from a small number of mutations and therefore have high rates of turnover. Furthermore, when IDRs contain multiple copies of a motif that jointly mediate a high-avidity interaction~\cite{Tompa2014} or a graded response to a signal via the accumulation of multiple phosphorylations~\cite{Wright2014}, the individual motifs are under weak selective constraints. As a result, though SLiMs are encoded by specific sequences, in some contexts they may evolve as distributed features that characterize IDRs as a whole rather than specific sites within them~\cite{Tan2010}.

@@

The GO annotation enrichment analysis indicates that proteins involved in DNA repair and extracellular structures contain a disproportionate number of regions whose disorder scores are rapidly evolving. Because the significance tests were not corrected for multiple testing or controlled for their false discovery rate, we caution against over-interpreting this result and instead consider it as a hypothesis for further investigation. In general, however, these regions with rapidly evolving disorder scores may correspond to molecular recognition features (MoRFs), which are modules in IDRs that undergo a disorder-to-order transition on binding their targets~\cite{Mohan2006, vanderLee2014}. Because MoRFs already exist on the boundary between disorder and structure, small changes in the biophysical properties of these regions may have disproportionate effects on their structural ensembles. Thus, the most variable disorder scores may reflect instances where a mutation triggered a ``phase transition'' between largely structured or disordered native states.

@@

\subsection{Brownian motion and Ornstein-Uhlenbeck analyses}
Brownian motion (BM) model parameters were calculated with two methods. The first used Felsenstein's contrasts algorithm to efficiently calculate roots and contrasts for the disorder scores and features of each region~\cite{Felsenstein1973, Felsenstein1985}. Rates were calculated as the mean of the squares of the contrasts. Though these values are unbiased, they are not maximum likelihood estimates and are inappropriate for use with the Akaike information criterion (AIC)~\cite{Akaike1974}. Thus, they were used for analyses involving only the BM model. For comparison with the Ornstein-Uhlenbeck (OU) model, the BM parameters were calculated by maximizing the likelihood. The OU model parameters were also calculated via maximum likelihood estimation as described in Butler \textit{et al.}~\cite{Butler2004}. To ensure parameter identfiability and estimate consistency, the root was treated as a random variable~\cite{Ho2013, Ho2014}. Thus, the covariance matrix, $V$, was parameterized as $V_{ij} = e^{\alpha d_{ij}}$ where $d_{ij}$ is the tree distance between tips $i$ and $j$, and $\alpha$ is the selection strength~\cite{Ho2013}.
